\OurTool{}: Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles\\[1em] Supplementary material

### 2019-06-20 IPANEMAP-2019.tex
Updates from Overleaf

\usetikzlibrary{shapes.geometric, arrows,calc,fit,positioning}

@@

\address{%
$^{1}$\,CNRS UMR 7161, LIX, Ecole Polytechnique, France
and
$^{2}$\,CitCoM, Cibles thérapeutiques et conception de Médicaments, CNRS, University Paris Descartes, 4 avenue de l'observatoire 75006 PARIS France}
% Affiliation must include:
% Department name, institution name, full road and district address,
% state, Zip or postal code, country

@@

\subsection*{The \OurTool{} method}
\bsi{Ce chapitre est vraiment très réussi, il explique vraiment bien ce qu'on fait et c'est accessible meme pour des polios de biologistes. je trouve dommage qu'il soit cantoné au Mat et Met. A mon avis il faudrait soit le mettre tel quel comme premier chapitre des résultats, soit en faire une version light sans les détails de calculs le mettre comme premier chapitre des résultats et ne laisser que les calculs dans le Mat et Met}  
We introduce the Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles (\OurTool{}), a novel approach that integrates the signal produced by various chemical probes. It performs clustering based on multiple sets of structures sampled in different experimental conditions, to ultimately return a set of structures that are representative of the dominant clusters. Its underlying rationale is that the prominent presence of a stable secondary structure within the  (pseudo-)Boltzmann ensembles induced by multiple experimental conditions should increase its likelihood to be (one of) the native structure(s) for a given RNA. It is thus hoped that integrating several reactivity profiles may be used to promote the native structure as one of the dominant structures within the multi-ensemble, and help circumvent the limitation of pseudo-energies derived from single reactivity profile, which are generally not sufficient to elect the native structure as its minimum (pseudo)-free energy candidate. In other words, combining ensembles of structures generated using multiple probing experiments is likely to denoise the Boltzmann (multi-)ensemble, and thus mitigate systematic biases induced by experimental conditions and reagents.

@@

\paragraph{Hadjin \emph{et al} dataset.} 
A dataset was gathered by Hadjin\etal\cite{Hajdin2013} to validate the predictive capacities of probing data-driven predictions. It consists of $24$ RNA sequences with known secondary structures for which a single chemical probing reactivity profile (SHAPE -- 1M7) is available.
This dataset includes sequences originating from a variety of organisms, and spans lengths ranging from $34$ nts to $530$ nts, with a focus on riboswitches and complex RNA architectures.

@@

\subsection*{Experimental probing protocols}
The structure of the Lariat capping ribozyme from \textit{Didymium iridis }(DiLCrz) was probed with different SHAPE reagents: 1M7 (1-methyl-7 nitrosatoic anhydre), NMIA (N-Methyl Isatoic Anhydre), BzCN (Benzoyl cyanide) and NAI (2-methylnicotinic acid imidazolide) in presence and absence of magnesium ions, and with DMS (DiMethylSulfate) and CMCT only in presence of magnesium ions.
As a preliminary experiment, we verified that DiLCrz generally adopts a single global architecture. To this end, we subjected DiLCrz to a standard denaturation protocol (80\degree{}C for 2 min in H2O, addition of 40 mM of HEPES at 7.5 pH, 100 mM of KCl, 5 mM of MgCl2, followed by 10 min at room temperature, and 10 min at 37\degree{}C), and observed the production of a single band on a non-denaturing PAGE, strongly suggesting the adoption of a single conformation.

\paragraph{Stops-inducing probing protocol (SHAPE-CE).} 
6 pmol of RNA were resuspended in 18 $\mu$l of water, denatured at 80\degree{}C for 2 minutes and cooled down at room temperature during 10 minutes in the probing buffer (40 mM HEPES at 7.5 pH, 100 mM KCl, in presence or absence of 5 mM MgCl2). After a 10 minutes incubation at 37\degree{}C, RNAs were treated with 2 mM of SHAPE reagent or DMSO (negative control) and incubated for 2 (BzCN), 5 (1M7), 30 (NMIA) or 60 (NAI) minutes at 37\degree{}C. Modified or unmodified RNAs were purified by ethanol precipitation and pellets were resuspended in 10 $\mu$l of water.
Modifications were revealed by reverse transcription using 5’ fluorescently labelled primers (D2 or D4 WellRED, Sigma Aldrich 5’ CTG-TGA-ACT-AAT-GCT-GTC-CTT-TAA 3’) and M-MLV RNAse H- reverse transcriptase (Promega) as previously described~\cite{Deforges2017}, with only minor modification of the originally described SHAPE protocol~\cite{Wilkinson2006}. cDNAs were separated by capillary electrophoresis (Beckman Coulter, Ceq8000). Data were analyzed using the  QuSHAPE~\cite{Karabiber2013} software. RNA probing was performed in triplicate with distinct RNA preparations.

@@

\paragraph{DiMethylSulfate (DMS) and CMCT probing.}
DMS and CMCT probing were conducted essentially as described in~\cite{Weill2004,James2008}. Succinctly, 6 pmol of RNA were resuspended in 18 $\mu$l of water, denatured at 80\degree{}C for 2 minutes and cooled down at room temperature for 10 minutes in a probing buffer for DMS probing (40 mM HEPES pH 7.5, 100 mM KCl, 5 mM MgCl2) or in CMCT (50 mM of potassium borate pH 8, 100 mM KCl and 5 mM MgCl2). For DMS, RNA was then treated with 1 $\mu$l of DMS (1:12 in ethanol) or 1 $\mu$l of ethanol for 5 minutes at 37\degree{}C (mock reaction), the reaction was stopped by addition of 400 mM of Tris at 7.5 pH and immediately put on ice. For CMCT,  RNA was treated with 10 $\mu$l of CMCT (42mg/mL) or 10 $\mu$l H2O for 10 minutes at 37\degree{}C (mock reaction). Modified RNA were ethanol-precipitated and resuspended in 10 $\mu$l of water. Modifications were mapped as described above for SHAPE-CE experiments.

@@

\begin{figure}
	\includegraphics[width=\linewidth]{graphs/RsampleVsIPANEMAP/Accuracy}
	\caption{Geometric mean of positive predictive value and sensitivity achieved by the predictions of \OurTool{} and \Software{Rsample} over the Hadjin\etal\cite{Hajdin2013} data set, consisting of 24 RNAs (RNA length indicated within square brackets). \label{fig:Vsrsample}}
\end{figure}

@@

	\caption{Model of the Lariat capping ribozyme from {\itshape Didymium iridis} (A -- Meyer\etal\cite{Meyer2014}), pseudoknot-free secondary structure as annotated by \Software{DSSR} (B -- pseudoknot ommitted) and \OurTool first-ranking prediction (C.) from 8 maximally-diverging probing conditions (NMIAMg, 1M7ILU3Mg, NAIMg, 1M7, CMCTMg,1M7ILU3,1M7Mg). \label{fig:didy}}
\end{figure*}
As benchmark RNA we used the lariat capping ribozyme from \textit{Didymium Iridis} (DiLCrz) which 3D structure has been modeled at 3,5 A from X-ray diffraction data~\cite{Meyer2014}. This RNA contains a large diversity of interactions, including helical domains, tertiary pairings, a kissing loop and a pseudoknot (Figure~\ref{fig:didy}).
It was probed as described in material and methods in 16 different conditions using different SHAPE and classical reagents, in presence/absence of Mg$^\text{2+}$ ions, using either the standard premature stop or the newly developed mutation induction technologies to map the modification sites. These experimental results were then used as input of \OurTool{}, either individually or combined,  for secondary structure prediction.

@@

	\caption{Discretized ensemble distance induced by conditions. Conditions ordered by spectral biclustering. Conditions regrouped within clusters by k-mean clustering appear as blocks on the x and y axis.}\label{fig:biclustering}
\end{figure}

@@

Unsurprisingly, the presence/absence of Magnesium during the probing impacts the predictions, with an observed drop from an average 73\% MCC in the presence of Mg$^\text{2+}$ to 67\% MCC in the absence of Mg$^\text{2+}$. However, while the poorly performing conditions of cluster \Bull{B} are probed in the absence of Mg$^\text{2+}$, some of the most informative profiles (\OneMSevILU and \NMIA) were obtained in absence of Mg$^\text{2+}$.
An even more drastic change of performances can be observed when comparing the MCC of mutation-based protocols (avg 78\%) with stop-based ones (avg 66\%).

@@

	%data are from multiprobing7 
	\caption{Comparison of predicted secondary structures for the \didy{} obtained with \OurTool{} and computational methods from a mono-probing experiment. Averaged MCC of structures predicted with \OurTool{} and with classic deterministic alternatives. The structure was probed in presence (\Mg) or absence (\NoMg) of Magnesium, using either stops (\Stop) or mutations (\Mut) inducing technologies.
	}
	\label{tab:thermo-ipan}
\end{table}

@@

\paragraph{Bi-probing mitigates the influence of outliers, but generally fails to improve predictions quality.}
We turn to a systematic exploration of the predictive capacities of bi-probing analyses, based on pairs of probing profiles, and attempt to quantify their impact on \OurTool predictions. For each pair and triplets of conditions, we executed \OurTool using 1~000 samples per condition, and considered the first returned structure. We then computed the associated MCC, and compare it with the MCC of the worst performing condition (Min), best-performing condition (Max) and average MCC over the single conditions experiments. A summary of the results over pairs is shown in Figure~\ref{fig:pairwise}.

A superficial inspection of the resulting MCCs appears to confirm the conclusions of Section~\ref{sec:cordero}. Indeed, considering two conditions leads to predictions whose quality fall between the worst and the best one, while being generally close to the average. More precisely, the average MCC value of pair-informed predictions remains at 70.7\%, improving only by 0.2\% over the average MCC of mono-probing predictions. Pair-informed predictions are marginally more reliable, with the median MCC increases to 72.5\% MCC from 70.5\% for mono-probing predictions. However, outliers strongly impact this picture, and ignoring the conditions in cluster \Bull{B} increases the average MCC to 75.1\%, while unsurprisingly decreasing the standard deviation (8.6\% $\to$ 6.5\%).

@@

\paragraph{Considering diverse conditions improves and stabilizes the quality of predictions.} 
Having established the redundant nature of similar conditions, and their limited contribution to downstream modeling, we turn to an analysis of conditions across clusters.  We run \OurTool{} the quality of predictions obtained by choosing a single condition in each of the 8 clusters. We consider any possible combination, except for the outlier cluster~\Bull{B}, for which we only consider \NAICE as a representative.

@@

For the WT, the average MCC achieved by our predictions is of 61.83\%, comparable to the worst condition of our previous dataset.
Our predictions are generally stable and, in 94\% out of the 100 runs, yield a structure having MCC between 60.5\% and 61\%. The remaining 6\% of the runs present improved MCC values, ranging from 76\% to 78\%, suggesting the existence of an alternative conformer in the pseudo-Boltzmann ensemble.
%%\as{ This value can not be fixed since we have  0.6134: 12\%, 0.7597: 2\%, 0.7649: 1\%, 0.7783: 1\%, 0.7665: 1\%, 0.7531: 1\% , is it ok Yann not to specify the improvemnt percentage ?}
%% Counter({0.6084: 82, 0.6134: 12, 0.7597: 2, 0.7649: 1, 0.7783: 1, 0.7665: 1, 0.7531: 1})
.%mean(Multiprobing_100WT)

@@

The average MCC again increases in the presence of 3 mutants, to reach an average of 69.3\% MCC. For 10 mutants, the average MCC and overall distribution remains highly similar to that obtained by considering three mutants, suggesting peeking performances. These results support a notion of complementarity between the probing data produced across reasonably similar mutants, leading to a gradual decrease of the noise in the data. However, the bimodal nature of the distribution also suggests the existence of at least two conformations that are both supported by the WT and the mutant profiles.

@@

To the best of our knowledge, \OurTool currently represents the only available method for a joint automated consideration of multiple probing profiles. Surprisingly, it also appears to produce the most accurate secondary structure models when informed by a single condition, as demonstrated by our reanalysis on the Hadjin\etal\cite{Hajdin2013} benchmark. We attribute this unexpected result to the adoption of a sampling/clustering approach, which seems to eliminate spurious outliers in the conformational space. In particular, the dominance of \OurTool is surprising since, while independently developed, our method shares the sampling/clustering philosophy of the recently introduced \Software{RSample}~\cite{Spasic2017}. We attribute these increased performances in part to a less stringent contribution of the pseudo-potentials, and to our adaptive choice of the number of clusters. Indeed, even in cases where a single dominant conformation is expected, considering multiple clusters seems to create more opportunities for discarding noisy random unstable conformations. Such conformations naturally occur in a Boltzmann-distributed subset of structures, and may end up \emph{polluting} the centroid structures by artificially inflating the cluster diversity. 
%The impact on \Software{RSample} predictions of such a \emph{noise} seems worsened by the fact that it returns the MEA, the centroid of the whole ensemble, whenever a dominant conformation is detected.

@@

A future work include solving the dilemma of whether or not to include Mg$^{\text{2+}}$ ions during the probing, in the perspective of a downstream computational prediction. On the one hand, RNA reaches its native 3D structure in the presence of Mg$^{\text{2+}}$ ions~\cite{Knapp1989}. However, such native structures include non WC pairings, and complex topological motifs such as pseudoknots, kissing loops, tetraloop interactions, that classic structure prediction algorithms do not explicitly capture. In contrast, in the presence of only monovalent ions, RNA is expected to form its helical domains, but leave its  tertiary contacts unstable. Therefore, probing the RNA without Mg$^{\text{2+}}$ could provide a signal that is less deceptive, and thus more informative, to predicting algorithms. Unfortunately, destabilizing tertiary contacts increases the probability of misfolding, or the coexistence in solution of multiple conformations. In this work, we probed a Didymium ribozyme which contains many tertiary motifs, systematically in the presence and absence of Mg$^{\text{2+}}$ (except for CMCT and DMS probes). Remarkably, the most accurate prediction was obtained with 1M7 and NMIA, irrespectively of the presence/absence of Mg$^{\text{2+}}$, this despite the fact that the reactivity profiles substantially diverged between the two conditions. However, for other reagents/protocols, predictions appeared positively impacted (+$10\%$ average MCC) by the presence of Mg$^{\text{2+}}$, painting a mixed picture that would deserve future investigations.
%\item je pense qu'il serait intéressant au moins pour un uo deux réactif de voir ce que donne le couple + et - Mg. Peut etre par exemple avec le 1M7ilu et un autre mauvais genre le NAI
%\end{itemize}
% il faudrait regrder ce que fait weeks avec le 1M7, le NMIA et le 1M6, je n'ai pas bien compris quelle tambouille il fait, mais ca part du meme genre de constatation : il y a une spécificité dans chacune des sondes qu'il n'arrive pas vraiment à formaliser donc il fait une tambouille qui lui permet de tous les intégrer (mais c'est peut etre un peu pédestre comme démarche!). Apres ce sont probablement des conditions tres proches ce qui n'apporte pas d'apres nos résultats bcq d'infos supplémentaires .. bref ca vaudrait peut etre le coup de comparer.

\begin{figure}
	{\centering\includegraphics[width=1\linewidth]{graphs/boxplotreproducibility.png}\\}

@@

\section*{Acknowledgements}
The authors wish to express their gratitude to Ronny Lorenz for developing a probing-informed version of \Software{RNAsubopt}, allowing to sample the pseudo-Boltzmann ensemble, and Nathalie Chamond for helpful suggestions.

>>>>

\usetikzlibrary{shapes.geometric, arrows, calc, fit, positioning}

@@

\address{%
$^{1}$\,CNRS UMR 7161, LIX, Ecole Polytechnique, France
and
$^{2}$\,CitCoM, Cibles thérapeutiques et conception de Médicaments, CNRS, Université de Paris, 4 avenue de l'observatoire 75006 PARIS France}
% Affiliation must include:
% Department name, institution name, full road and district address,
% state, Zip or postal code, country

@@

\subsection*{The \OurTool{} method}
\bsi{Ce chapitre est vraiment très réussi, il explique vraiment bien ce qu'on fait et c'est accessible meme pour des polios de biologistes. je trouve dommage qu'il soit cantoné au Mat et Met. A mon avis il faudrait soit le mettre tel quel comme premier chapitre des résultats, soit en faire une version light sans les détails de calculs le mettre comme premier chapitre des résultats et ne laisser que les calculs dans le Mat et Met}  
\ypi{Dans la mesure où l'intro est courte, et étant donné le titre, je ne pense pas que sa présence en Mat Met limite  sur visibilité. En plus, c'est typiquement là où tu le trouves d'habitude dans ce type de papier (cf Rsample).}
We introduce the Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles (\OurTool{}), a novel approach that integrates the signal produced by various chemical probes. It performs clustering based on multiple sets of structures sampled in different experimental conditions, to ultimately return a set of structures that are representative of the dominant clusters. Its underlying rationale is that the prominent presence of a stable secondary structure within the  (pseudo-)Boltzmann ensembles induced by multiple experimental conditions should increase its likelihood to be (one of) the native structure(s) for a given RNA. It is thus hoped that integrating several reactivity profiles may be used to promote the native structure as one of the dominant structures within the multi-ensemble, and help circumvent the limitation of pseudo-energies derived from single reactivity profile, which are generally not sufficient to elect the native structure as its minimum (pseudo)-free energy candidate. In other words, combining ensembles of structures generated using multiple probing experiments is likely to denoise the Boltzmann (multi-)ensemble, and thus mitigate systematic biases induced by experimental conditions and reagents.

@@

\paragraph{Hajdin \emph{et al} dataset.} 
A dataset was gathered by Hajdin\etal\cite{Hajdin2013} to validate the predictive capacities of probing data-driven predictions. It consists of $24$ RNA sequences with known secondary structures for which a single chemical probing reactivity profile (SHAPE -- 1M7) is available.
This dataset includes sequences originating from a variety of organisms, and spans lengths ranging from $34$ nts to $530$ nts, with a focus on riboswitches and complex RNA architectures.

@@

\subsection*{Experimental probing protocols}
The structure of the Lariat capping ribozyme from \textit{Didymium iridis }(DiLCrz) was probed with different SHAPE reagents: 1M7 (1-methyl-7 nitrosatoic anhydre), NMIA (N-Methyl Isatoic Anhydre), BzCN (Benzoyl cyanide) and NAI (2-methylnicotinic acid imidazolide) in presence and absence of Mg$^\text{2+}$, and with DMS (DiMethylSulfate) and CMCT only in presence of Mg$^\text{2+}$.
As a preliminary experiment, we verified that DiLCrz generally adopts a single global architecture. To this end, we subjected DiLCrz to a standard denaturation protocol (80\degree{}C for 2 min in H$_\text{2}$O, addition of 40 mM of HEPES at 7.5 pH, 100 mM of KCl, 5 mM of MgCl$_\text{2}$, followed by 10 min at room temperature, and 10 min at 37\degree{}C), and observed the production of a single band on a non-denaturing PAGE, strongly suggesting the adoption of a single conformation.

\paragraph{Stops-inducing probing protocol (SHAPE-CE).} 
6 pmol of RNA were resuspended in 18 $\mu$l of water, denatured at 80\degree{}C for 2 minutes and cooled down at room temperature during 10 minutes in the probing buffer (40 mM HEPES at 7.5 pH, 100 mM KCl, in presence or absence of 5 mM MgCl$_\text{2}$. After a 10 minutes incubation at 37\degree{}C, RNAs were treated with 2 mM of SHAPE reagent or DMSO (negative control) and incubated for 2 (BzCN), 5 (1M7), 30 (NMIA) or 60 (NAI) minutes at 37\degree{}C. Modified or unmodified RNAs were purified by ethanol precipitation and pellets were resuspended in 10 $\mu$l of water.
Modifications were revealed by reverse transcription using 5’ fluorescently labelled primers (D2 or D4 WellRED, Sigma Aldrich 5’ CTG-TGA-ACT-AAT-GCT-GTC-CTT-TAA 3’) and M-MLV RNAse H- reverse transcriptase (Promega) as previously described~\cite{Deforges2017}, with only minor modification of the originally described SHAPE protocol~\cite{Wilkinson2006}. cDNAs were separated by capillary electrophoresis (Beckman Coulter, Ceq8000). Data were analyzed using the  QuSHAPE~\cite{Karabiber2013} software. RNA probing was performed in triplicate with distinct RNA preparations.

@@

\paragraph{DiMethylSulfate (DMS) and CMCT probing.}
DMS and CMCT probing were conducted essentially as described in~\cite{Weill2004,James2008}. Succinctly, 6 pmol of RNA were resuspended in 18 $\mu$l of water, denatured at 80\degree{}C for 2 minutes and cooled down at room temperature for 10 minutes in a probing buffer for DMS probing (40 mM HEPES pH 7.5, 100 mM KCl, 5 mM MgCl$_\text{2}$) or in CMCT (50 mM of potassium borate pH 8, 100 mM KCl and 5 mM MgCl$_\text{2}$). For DMS, RNA was then treated with 1 $\mu$l of DMS (1:12 in ethanol) or 1 $\mu$l of ethanol for 5 minutes at 37\degree{}C (mock reaction), the reaction was stopped by addition of 400 mM of Tris at 7.5 pH and immediately put on ice. For CMCT,  RNA was treated with 10 $\mu$l of CMCT (42mg/mL) or 10 $\mu$l H$_\text{2}$O for 10 minutes at 37\degree{}C (mock reaction). Modified RNA were ethanol-precipitated and resuspended in 10 $\mu$l of water. Modifications were mapped as described above for SHAPE-CE experiments.

@@

\begin{figure}
	\includegraphics[width=\linewidth]{graphs/RsampleVsIPANEMAP/Accuracy}
	\caption{Geometric mean of positive predictive value and sensitivity achieved by the predictions of \OurTool{} and \Software{Rsample} over the Hajdin\etal\cite{Hajdin2013} data set, consisting of 24 RNAs (RNA length indicated within square brackets). \label{fig:Vsrsample}}
\end{figure}

@@

	\caption{Model of the Lariat capping ribozyme from {\itshape Didymium iridis} (A -- Meyer\etal\cite{Meyer2014}), pseudoknot-free secondary structure as annotated by \Software{DSSR} (B -- pseudoknot ommitted) and \OurTool first-ranking prediction (C.) from 8 maximally-diverging probing conditions (NMIAMg, 1M7ILU3Mg, NAIMg, 1M7, CMCTMg,1M7ILU3,1M7Mg). \bsi{figure à refaite. Faut il montrer les interactions tertiaires? } \label{fig:didy}}
\end{figure*}
As benchmark RNA we used the lariat capping ribozyme from \textit{Didymium Iridis} (DiLCrz) which 3D structure has been modeled at 3,5 A from X-ray diffraction data~\cite{Meyer2014}. This RNA contains a large diversity of interactions, including helical domains, tertiary pairings, a kissing loop and a pseudoknot (Figure~\ref{fig:didy}).
It was probed as described in material and methods in 16 different conditions using different SHAPE and classical reagents, in presence/absence of Mg$^\text{2+}$ ions, using either the standard premature stop or the newly developed mutation induction technologies to map the modification sites. These experimental results were then used as input of \OurTool{}, either individually or combined,  for secondary structure prediction.

@@

	\caption{Discretized ensemble distance induced by conditions. Conditions ordered by spectral biclustering. Conditions regrouped within clusters by k-mean clustering appear as blocks on the x and y axis.\bsi{pourquoi les clusters E et F ne sont ils pas regroupés comme les autres?}}\label{fig:biclustering}
\end{figure}

@@

Unsurprisingly, the presence/absence of Magnesium ions during the probing impacts the predictions, with an observed drop from an average 73\% MCC in the presence of Mg$^\text{2+}$ to 67\% MCC in the absence of Mg$^\text{2+}$. However, while the poorly performing conditions of cluster \Bull{B} are probed in the absence of Mg$^\text{2+}$, some of the most informative profiles (\OneMSevILU and \NMIA) were obtained in absence of Mg$^\text{2+}$.
An even more drastic change of performances can be observed when comparing the MCC of mutation-based protocols (avg 78\%) with stop-based ones (avg 66\%).

@@

	%data are from multiprobing7 
	\caption{Comparison of predicted secondary structures for the \didy{} obtained with \OurTool{} and computational methods from a mono-probing experiment. Averaged MCC of structures predicted with \OurTool{} and with classic deterministic alternatives. The structure was probed in presence (\Mg) or absence (\NoMg) of Magnesium ions, using either stops (\Stop) or mutations (\Mut) inducing technologies.
	}
	\label{tab:thermo-ipan}
\end{table}

@@

\paragraph{Bi-probing mitigates the influence of outliers, but generally fails to improve predictions quality.}
We turn to a systematic exploration of the predictive capacities of bi-probing analyses, based on pairs of probing profiles, and attempt to quantify their impact on \OurTool predictions. For each pair and triplets of conditions, we executed \OurTool using 1000 samples per condition, and considered the first returned structure. We then computed the associated MCC, and compare it with the MCC of the worst performing condition (Min), best-performing condition (Max) and average MCC over the single conditions experiments. A summary of the results over pairs is shown in Figure~\ref{fig:pairwise}.

A superficial inspection of the resulting MCCs appears to confirm the conclusions of Section Bruno~\ref{sec:cordero}. Indeed, considering two conditions leads to predictions whose quality fall between the worst and the best one, while being generally close to the average. More precisely, the average MCC value of pair-informed predictions remains at 70.7\%, improving only by 0.2\% over the average MCC of mono-probing predictions. Pair-informed predictions are marginally more reliable, with the median MCC increases to 72.5\% MCC from 70.5\% for mono-probing predictions. However, outliers strongly impact this picture, and ignoring the conditions in cluster \Bull{B} increases the average MCC to 75.1\%, while unsurprisingly decreasing the standard deviation (8.6\% $\to$ 6.5\%).

@@

\paragraph{Considering diverse conditions improves and stabilizes the quality of predictions.} 
Having established the redundant nature of similar conditions, and their limited contribution to downstream modeling, we turn to an analysis of conditions across clusters.  We run \OurTool{} the quality of predictions obtained by choosing a single condition in each of the 8 clusters.\bsi{il manque des trucs dans cette phrase (celle qui commence par we run), non?} We consider any possible combination, except for the outlier cluster~\Bull{B}, for which we only consider \NAICE as a representative.

@@

For the WT, the average MCC achieved by our predictions is of 61.83\%, comparable to the worst condition of our previous dataset.
Our predictions are generally stable and, in 94\% out of the 100 runs, yield a structure having MCC between 60.5\% and 61\%. The remaining 6\% of the runs present improved MCC values, ranging from 76\% to 78\%, suggesting the existence of an alternative conformer in the pseudo-Boltzmann ensemble.
%%\as{ This value can not be fixed since we have  0.6134: 12\%, 0.7597: 2\%, 0.7649: 1\%, 0.7783: 1\%, 0.7665: 1\%, 0.7531: 1\% , is it ok Yann not to specify the improvemnt percentage ?}
%% Counter({0.6084: 82, 0.6134: 12, 0.7597: 2, 0.7649: 1, 0.7783: 1, 0.7665: 1, 0.7531: 1})
%mean(Multiprobing_100WT)

@@

The average MCC again increases in the presence of 3 mutants, to reach an average of 69.3\% MCC. For 10 mutants, the average MCC and overall distribution remains highly similar to that obtained by considering three mutants, suggesting peeking performances. These results support a notion of complementarity between the probing data produced across reasonably similar mutants, leading to a gradual decrease of the noise in the data. However, the bimodal nature of the distribution also suggests the existence of at least two conformations that are both supported by the WT and the mutant profiles.
\bsi{il y a une conformation alternative connue. Tu pourrais me donner un model provenant des 2 clusters de MCC que je regarde si pas hasard c'est ca?}

@@

To the best of our knowledge, \OurTool currently represents the only available method for a joint automated consideration of multiple probing profiles. Surprisingly, it also appears to produce the most accurate secondary structure models when informed by a single condition, as demonstrated by our reanalysis on the Hajdin\etal\cite{Hajdin2013} benchmark. We attribute this unexpected result to the adoption of a sampling/clustering approach, which seems to eliminate spurious outliers in the conformational space. In particular, the dominance of \OurTool is surprising since, while independently developed, our method shares the sampling/clustering philosophy of the recently introduced \Software{RSample}~\cite{Spasic2017}. We attribute these increased performances in part to a less stringent contribution of the pseudo-potentials, and to our adaptive choice of the number of clusters. Indeed, even in cases where a single dominant conformation is expected, considering multiple clusters seems to create more opportunities for discarding noisy random unstable conformations. Such conformations naturally occur in a Boltzmann-distributed subset of structures, and may end up \emph{polluting} the centroid structures by artificially inflating the cluster diversity. 
%The impact on \Software{RSample} predictions of such a \emph{noise} seems worsened by the fact that it returns the MEA, the centroid of the whole ensemble, whenever a dominant conformation is detected.

@@

A future work include solving the dilemma of whether or not to include Mg$^{\text{2+}}$ ions during the probing, in the perspective of a downstream computational prediction. On the one hand, RNA reaches its native 3D structure in the presence of Mg$^{\text{2+}}$ ions~\cite{Knapp1989}. However, such native structures include non WC pairings, and complex topological motifs such as pseudoknots, kissing loops, tetraloop interactions, that classic structure prediction algorithms do not explicitly capture. In contrast, in the presence of only monovalent ions, RNA is expected to form its helical domains, but leave its  tertiary contacts unstable, with the notable exception of the G quadruplex. Therefore, probing the RNA without Mg$^{\text{2+}}$ could provide a signal that is less deceptive, and thus more informative, to predicting algorithms. Unfortunately, destabilizing tertiary contacts increases the probability of misfolding, or the coexistence in solution of multiple conformations. In this work, we probed a Didymium ribozyme which contains many tertiary motifs, systematically in the presence and absence of Mg$^{\text{2+}}$ (except for CMCT and DMS probes). Remarkably, the most accurate prediction was obtained with 1M7 and NMIA, irrespectively of the presence/absence of Mg$^{\text{2+}}$, this despite the fact that the reactivity profiles substantially diverged between the two conditions. However, for other reagents/protocols, predictions appeared positively impacted (+$10\%$ average MCC) by the presence of Mg$^{\text{2+}}$, painting a mixed picture that would deserve future investigations.
%\item je pense qu'il serait intéressant au moins pour un uo deux réactif de voir ce que donne le couple + et - Mg. Peut etre par exemple avec le 1M7ilu et un autre mauvais genre le NAI
%\end{itemize}
% il faudrait regrder ce que fait weeks avec le 1M7, le NMIA et le 1M6, je n'ai pas bien compris quelle tambouille il fait, mais ca part du meme genre de constatation : il y a une spécificité dans chacune des sondes qu'il n'arrive pas vraiment à formaliser donc il fait une tambouille qui lui permet de tous les intégrer (mais c'est peut etre un peu pédestre comme démarche!). Apres ce sont probablement des conditions tres proches ce qui n'apporte pas d'apres nos résultats bcq d'infos supplémentaires .. bref ca vaudrait peut etre le coup de comparer.

\begin{figure}
	{\centering\includegraphics[width=1.2\linewidth]{graphs/boxplotreproducibility.png}\\}

@@

\section*{Acknowledgements}
The authors wish to express their gratitude to Ronny Lorenz for developing a probing-informed version of \Software{RNAsubopt}, allowing to sample the pseudo-Boltzmann ensemble, Nathalie Chamond for helpful suggestions and Benoit Masquida for the gift of DiLCRz plasmid and helpful discussions about protocols and DiLCRz structure .

### 2019-06-20 SuppMat.tex
Updates from Overleaf

{\centering
	\begin{adjustbox}{max width=\textwidth,max height=\textheight,keepaspectratio}
\begin{tabular}{@{}c|cccc@{}}

%\toprule
 \includegraphics[scale=\MyScale]{graphs/Supp_structures/native_structure}&
 \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILU_NMIAMg}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILUMg_1M7ILU3Mg}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7MgCE_NMIAMgCE}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/NaiMg_CMCTMg}\\
 \midrule
Native&\OneMSevILU&\OneMSevILUMg&\OneMSevMgCE&\NAIMg\\
structure &+\NMIAMg&+\OneMSevILUThreeMg&+\NMIAMgCE&+\CMCTMg\\
\end{tabular}
\end{adjustbox}\\}

>>>>

{\centering
	\begin{adjustbox}{max width=\textwidth,max height=\textheight,keepaspectratio}
\begin{tabular}{@{}c|ccccc@{}}

%\toprule
 \includegraphics[scale=\MyScale]{graphs/Supp_structures/native_structure}&
 \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILU_NMIAMg}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILUMg_1M7ILU3Mg}&
 \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7MgCE_NMIAMgCE}
 &
 \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7CE_NMIACE}
 & \includegraphics[scale=\MyScale]{graphs/Supp_structures/NaiMg_CMCTMg}\\
 \midrule
Native&\OneMSevILU&\OneMSevILUMg&\OneMSevMgCE& \1M7CE &\NAIMg\\
structure &+\NMIAMg&+\OneMSevILUThreeMg&+\NMIAMgCE&+\NMIACE &+\CMCTMg\\
\end{tabular}
\end{adjustbox}\\}
